---
title: "Mojave_Div_ITS_phyloseq"
author: "Nat Pombubpa"
date: "Updated on December 11, 2019"
output: html_document
---

#Fungal taxonomic barplot by site

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###STEP1: Load all necessary packages for analysis
More information about Phyloseq can be found at the following link: [Phyloseq](https://joey711.github.io/phyloseq/)
If you get error in this step, you probably need to install any packages which causes error.

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(plyr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(ggtree)
library(treeio)
library(devtools)
library(MicEco)
library(MoMAColors)
library(castor)


```

###STEP2: Import Mapping file (metadate file)
1.Check mapping file before import to R, R doesn’t seem to like sample name to start with number or contain “-” in sample name. If you get error in this step, you should check file name first.

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating "KEEP" in Description column

```{r}
meta = read.table("Metadata_KHK2.tsv",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

If need to keep or remove some samples
```{r}
#meta <- meta[which(meta$Layer %in% c("Surface")),]
```

###STEP3: Check if your metadata file has been import successfully and correctly

The output will show a table of your metadata file (mapping file).

*If you do not have header, you might start your first row with #

```{r warning=FALSE}
head(meta)
```

###STEP4: Construct sample_data-class using imported metadata

```{r}
sampleData <- sample_data(meta)
```

###STEP5: Import OTU table

OTU table from Mojave 16S data is “MC2017FC.otu_table.txt”.
MC-Mojave Crust-2017-FC-Fungi completed

```{r}
otus <- read.table("Mojave2020_KKelly202307.otu_table.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

Check imported OTU table

```{r warning=FALSE}
head(OTU)
```

###STEP6: Import taxonomy table
Taxonmy table generated from AMPtk need to be rearranged using following script.

“perl rdp_taxonmy2mat.pl<Input_taxonmy.txt>Output_taxonomy.txt”

rdp_taxonomy2mat.pl was created by Professor Jason E. Stajich

```{r}
taxmat <- read.table("Mojave2020_KKelly202307.ASVs.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP7: Import phylogenetic tree
Phylogenetic tree can also be include for further phylogenetic analysis.

```{r warning=FALSE}
treefile = "Mojave2020_KKelly202307.ASVs.tree.phy"
tree = read.tree(treefile)
```

###STEP8: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData,tree)
```

Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object""

```{r warning=FALSE}
physeq
```

###STEP9: Remove singletons
Remove any OTUs that present only one time.

```{r }
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
```

```{r warning=FALSE}
physeq.prune
```

###STEP10: Plot read counts to check dataset
Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq.prune), "data.frame"),
                 TotalReads = sample_sums(physeq.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")

#For plotting, use command below.
#SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
head(readcount)
```

```{r warning=FALSE}
set.seed(1)
physeq.prune.rarefy = rarefy_even_depth(physeq.prune, sample.size = 7909, replace = FALSE, trimOTUs = FALSE)
physeq.prune.rarefy
```

###STEP11: Amphinema trees

```{r}
# Subset taxa for Glomus and Rhizoglomus
#physeq.prune.rarefy.Amphinemas <- subset_taxa(physeq.prune.rarefy, Genus %in% c("Glomus", "Rhizophagus"))
physeq.prune.rarefy.Amphinemas <- subset_taxa(physeq.prune.rarefy, Genus == "Amphinema")
physeq.prune.rarefy.Amphinemas

```


```{r}

#tax_table(physeq.prune.rarefy.Amphinemas)

```

```{r}

phy_tree(physeq.prune.rarefy.Amphinemas)

# Plot the modified tree
p2 <- ggtree(phy_tree(physeq.prune.rarefy.Amphinemas), layout="rectangular") +
  geom_tiplab(size=2, color="black")





#p2

```


```{r}
Dali <- moma.colors("Dali")

# Generate a larger palette of colors
#dali_palette <- moma.colors("Dali", n = 200)

# Subset the palette to include 100 colors
#palette_100 <- dali_palette[1:100]

# Display the first few colors in the palette
#head(palette_100)
```



```{r}

tax_table_prune <- tax_table(physeq.prune.rarefy.Amphinemas)
#tax_table_prune

# Extract species names from the taxonomy table
species_names <- tax_table_prune[, "Species"]
genus_names <-  tax_table_prune[, "Genus"]

species_names[is.na(species_names)] <- tax_table_prune[, "Genus"]

# If Species column is not available, you can extract Genus instead
#if (!"Species" %in% colnames(tax_table_prune)) {
#  species_names <- tax_table_prune[, "Genus"]
#}

species_names[is.na(species_names)] <- "Unknown Amphinema"

#species_names


```

```{r}

# Get the tip labels (ASV IDs) from the phylogenetic tree
tip_labels <- phy_tree(physeq.prune.rarefy.Amphinemas)$tip.label

# Match ASV IDs with species names and replace tip labels
new_tip_labels <- species_names[match(tip_labels, rownames(species_names))]

#new_tip_labels
```


```{r}

object <- phy_tree(physeq.prune.rarefy.Amphinemas)

tiplabels <- object$tip.label

new_tip_labels <- paste(tiplabels, new_tip_labels, sep = " ")

object$tip.label <- new_tip_labels

object
```

```{r}

# Plot the modified tree with species names as tip labels
p3 <- ggtree(object, layout="rectangular")#+
#  geom_tiplab(size=2, color="black")


```

```{r}

# Define the number of colors you want in the palette
num_colors <- 100

# Generate the palette from white to red
palette_white_to_red <- colorRampPalette(c("aquamarine4", "yellow1", "darkorange"))(num_colors)

# Display the first few colors in the palette
head(palette_white_to_red)
```

###STEP12: Plot tree with heatmap

```{r}
# Get the OTU table
otu_table <- otu_table(physeq.prune.rarefy.Amphinemas)

species_names[is.na(species_names)] <- "Unknown Amphinema"

# Replace row names in the OTU table with species names
rownames(otu_table) <- new_tip_labels

# Get the sample data from the phyloseq object
sample_data <- as.data.frame(sample_data(physeq.prune.rarefy.Amphinemas))

# Extract the variables you want to use as column names (e.g., soil type, site)
column_names <- paste(sample_data$Crust_type, sample_data$Site, sep = " ")



log_otu_table <- log10(otu_table + 1)  # Adding 1 to avoid log(0)

# Replace column names in the OTU table
colnames(log_otu_table) <- column_names

colnames(log_otu_table) <- make.unique(colnames(log_otu_table))


p_heatmap <- gheatmap(p3, log_otu_table, width=4, offset=0.25, 
                      legend_title="Relative Abundance", colnames_position="bottom", 
                      colnames_angle=45, colnames_offset_y=0, hjust=1, font.size=1.25) +
  scale_fill_gradientn("Relative abundance", colours = palette_white_to_red) +
  scale_y_continuous(limits=c(-2, 25)) + theme_tree(legend.position="right")
plot(p_heatmap)
```

```{r}

# Save the heatmap as a file
ggsave("Amphinema.png", plot = p_heatmap, width = 8, height = 6)

```


###STEP12: Plot tree with faceted bar plots


```{r}
# Get the OTU table
otu_table <- otu_table(physeq.prune.rarefy.Amphinemas)

# Replace row names in the OTU table with species names
rownames(otu_table) <- new_tip_labels

# Get the sample data from the phyloseq object
sample_data <- as.data.frame(sample_data(physeq.prune.rarefy.Amphinemas))

# Extract the variables you want to use as column names (e.g., soil type, site)
column_names <- paste(sample_data$Crust_type, sample_data$Site, sep = " ")

log_otu_table <- log10(otu_table + 1)  # Adding 1 to avoid log(0)

# Replace column names in the OTU table
colnames(log_otu_table) <- column_names

colnames(log_otu_table) <- make.unique(colnames(log_otu_table))


p_heatmap <- gheatmap(p3, log_otu_table, width=4, offset=0.25, 
                      legend_title="Relative Abundance", colnames_position="bottom", 
                      colnames_angle=45, colnames_offset_y=0, hjust=1, font.size=1.25) +
  scale_fill_gradientn("Relative abundance", colours = palette_white_to_red) +
  scale_y_continuous(limits=c(-2, 25)) + theme_tree(legend.position="right")
plot(p_heatmap)
```



```{r}
physeq.prune.rarefy.AmphinemasAB <- subset_samples(physeq.prune.rarefy.Amphinemas, Site == "AB")

physeq.prune.rarefy.AmphinemasODLO <- subset_samples(physeq.prune.rarefy.Amphinemas, Site == "ODLO")

physeq.prune.rarefy.AmphinemasTP <- subset_samples(physeq.prune.rarefy.Amphinemas, Site == "TP")

physeq.prune.rarefy.AmphinemasCIMA <- subset_samples(physeq.prune.rarefy.Amphinemas, Site == "CIMA")

physeq.prune.rarefy.AmphinemasGMT <- subset_samples(physeq.prune.rarefy.Amphinemas, Site == "GMT")

#sample_data <- as.data.frame(sample_data(physeq.prune.rarefy.Amphinemas))

AB = taxa_sums(physeq.prune.rarefy.AmphinemasAB)
#sums.Amphinemas

TP = taxa_sums(physeq.prune.rarefy.AmphinemasTP)
#sums.Amphinemas

GMT = taxa_sums(physeq.prune.rarefy.AmphinemasGMT)
#sums.Amphinemas

CIMA = taxa_sums(physeq.prune.rarefy.AmphinemasCIMA)
#sums.Amphinemas

ODLO = taxa_sums(physeq.prune.rarefy.AmphinemasODLO)
#sums.Amphinemas



AB <- data.frame(ID=new_tip_labels, AB)

AB <- AB %>%
  mutate(Site = "AB")

colnames(AB)[colnames(AB) == "AB"] <- "Count"


GMT <- data.frame(ID=new_tip_labels, GMT)

GMT <- GMT %>%
  mutate(Site = "GMT")

colnames(GMT)[colnames(GMT) == "GMT"] <- "Count"


TP <- data.frame(ID=new_tip_labels, TP)

TP <- AB %>%
  mutate(Site = "TP")

colnames(GMT)[colnames(GMT) == "TP"] <- "Count"


ODLO <- data.frame(ID=new_tip_labels, ODLO)

ODLO <- AB %>%
  mutate(Site = "ODLO")

colnames(ODLO)[colnames(ODLO) == "ODLO"] <- "Count"


CIMA <- data.frame(ID=new_tip_labels, CIMA)

CIMA <- CIMA %>%
  mutate(Site = "CIMA")

colnames(CIMA)[colnames(CIMA) == "CIMA"] <- "Count"


AB <- cbind(AB, new_tip_labels)


GMT <- cbind(GMT, new_tip_labels)


TP <- cbind(TP, new_tip_labels)


CIMA <- cbind(CIMA, new_tip_labels)

ODLO <- cbind(ODLO, new_tip_labels)


```

```{r}

b <- rbind(AB, GMT, TP, CIMA, ODLO)

# Define a small constant
epsilon <- 1  # You can adjust this value based on your preference

# Assuming your DataFrame is called df and the column is called Val
b$Count <- log10(b$Count + epsilon)

#b


total_sum <- d %>%
  filter(ID == "ASV980 Glomus macrocarpum") %>%
  summarise(total_sum = sum(Val, na.rm = TRUE))

print(total_sum$total_sum)


```
```{r}
#p4 <- facet_plot(p3, panel='Stacked Barplot', data=b, geom=geom_barh, aes(x=Site), color="firebrick") + theme_tree2()
#p4

p4 <- facet_plot(p3, panel = 'Site', data = b, 
				geom = geom_barh, 
				mapping = aes(x = Count, fill = Site), 
				stat='identity' ) 
p4



```



```{r}

# Save the heatmap as a file
ggsave("Amphinema_plot.png", plot = p4, width = 8, height = 6)

```


```{r}
physeq.prune.rarefy.AmphinemasSMC <- subset_samples(physeq.prune.rarefy.Amphinemas, Crust_type == "SMC")

physeq.prune.rarefy.AmphinemasSMC.ster <- subset_samples(physeq.prune.rarefy.Amphinemas, Crust_type == "SMC.ster")

physeq.prune.rarefy.AmphinemasDirt <- subset_samples(physeq.prune.rarefy.Amphinemas, Crust_type == "Dirt")

physeq.prune.rarefy.AmphinemasSMC.sub <- subset_samples(physeq.prune.rarefy.Amphinemas, Crust_type == "SMC.sub")


#sample_data <- as.data.frame(sample_data(physeq.prune.rarefy.Amphinemas))

SMC = taxa_sums(physeq.prune.rarefy.AmphinemasSMC)
#sums.Amphinemas

Dirt = taxa_sums(physeq.prune.rarefy.AmphinemasDirt)
#sums.Amphinemas

SMC.ster = taxa_sums(physeq.prune.rarefy.AmphinemasSMC.ster)
#sums.Amphinemas

SMC.sub = taxa_sums(physeq.prune.rarefy.AmphinemasSMC.sub)
#sums.Amphinemas
```



```{r}

SMC <- data.frame(ID=new_tip_labels, SMC)

SMC <- SMC %>%
  mutate(Crust_type = "SMC")

colnames(SMC)[colnames(SMC) == "SMC"] <- "Val"


SMC.ster <- data.frame(ID=new_tip_labels, SMC.ster)

SMC.ster <- SMC.ster %>%
  mutate(Crust_type = "SMC.ster")

colnames(SMC.ster)[colnames(SMC.ster) == "SMC.ster"] <- "Val"


SMC.sub <- data.frame(ID=new_tip_labels, SMC.sub)

SMC.sub <- SMC.sub %>%
  mutate(Crust_type = "SMC.sub")

colnames(SMC.sub)[colnames(SMC.sub) == "SMC.sub"] <- "Val"


Dirt <- data.frame(ID=new_tip_labels, Dirt)

Dirt <- Dirt %>%
  mutate(Crust_type = "Dirt")

colnames(Dirt)[colnames(Dirt) == "Dirt"] <- "Val"



Dirt <- cbind(Dirt, new_tip_labels)


SMC <- cbind(SMC, new_tip_labels)


SMC.ster <- cbind(SMC.ster, new_tip_labels)


SMC.sub <- cbind(SMC.sub, new_tip_labels)



```



```{r}

d <- rbind(Dirt, SMC, SMC.ster, SMC.sub)

# Define a small constant
epsilon <- 1  # You can adjust this value based on your preference

d$Val <- log10(d$Val + epsilon)

d

total_sum <- d %>%
  filter(ID == "ASV980 Glomus macrocarpum") %>%
  summarise(total_sum = sum(Val, na.rm = TRUE))

print(total_sum$total_sum)

```

```{r}
p5 <- facet_plot(p4, panel = 'Crust type', data = d, 
				geom = geom_barh, 
				mapping = aes(x = Val, fill = Crust_type), 
				stat='identity' ) +
  scale_fill_manual(values = Dali)
p5
```



```{r}

# Save the heatmap as a file
ggsave("Amphinemabarplot.png", plot = p5, width = 8, height = 6)

```


```{r}
p5 <- facet_plot(p4, panel = 'Crust type', data = d, 
				geom = geom_barh, 
				mapping = aes(x = Val, fill = Crust_type), 
				stat='identity' ) +
  scale_fill_manual(values = Dali)
p5
```

```{r}

library(castor)
```

```{r}
p5 <- facet_plot(p4, panel = 'Crust type', data = Dirt, 
				geom = geom_violinh, 
				mapping = aes(x = Val, fill = Crust_type), 
				stat='identity' ) +
  scale_fill_manual(values = Dali) 
#p5

#facet_plot(p4, panel="Abundance", data=Dirt, 
#           geom_joy, mapping = aes(x=Val,group=label, 
#                                        fill=Crust_type), 
#           color='grey80', lwd=.3)

#p4 <- facet_plot(p3, panel='Stacked Barplot', data=b, geom=geom_barh, aes(x=Site), color="firebrick") + theme_tree2()
#p4

p10 <- facet_plot(p3, panel = 'Site', data = Dirt, 
				geom = geom_barh, 
				mapping = aes(x = Val, fill = Crust_type), 
				stat='identity' ) 
epsilon <- 100  # You can adjust this value based on your preference

# Assuming your DataFrame is called df and the column is called Val
Dirt$log_Val <- Dirt$Val + epsilon


epsilon <- 1  # You can adjust this value based on your preference

# Assuming your DataFrame is called df and the column is called Val
Dirt$Val <- Dirt$Val + epsilon

facet_plot(p3, panel="Abundance", data=Dirt, 
           geom_joy, mapping = aes(x=Val,group=label, 
                                        fill=Crust_type), 
           color='grey80', lwd=.3)



#p4 <- facet_plot(p3, panel='Stacked Barplot', data=b, geom=geom_barh, aes(x=Site), color="firebrick") + theme_tree2()
#p4
           
```



```{r}
d
           
```

```{r}
SMC
           
```


```{r}

phy_tree(physeq.prune.rarefy.Amphinemas)

# Plot the modified tree
p100 <- ggtree(phy_tree(physeq.prune.rarefy.Amphinemas), layout="rectangular") +
  geom_tiplab(size=2, color="black")

labels <- p100$tip.label
#labels
d1 <- data.frame(id=p10$tip.label, location=sample(c(labels), 30, replace=TRUE))
new_tip_labels
#p1 <- p %<+% d1 + geom_tippoint(aes(color=location))
```

```{r}

# Get the OTU table
#otu_table <- otu_table(physeq.prune.rarefy.Amphinemas)

#species_names[is.na(species_names)] <- "Unknown Amphinema"

# Replace row names in the OTU table with species names
#rownames(otu_table) <- new_tip_labels

# Get the sample data from the phyloseq object
#sample_data <- as.data.frame(sample_data(physeq.prune.rarefy.Amphinemas))

# Extract the variables you want to use as column names (e.g., soil type, site)
#column_names <- paste(d$Crust_type, d$d$Crust_type, sep = " ")



log_otu_table <- log10(otu_table + 1)  # Adding 1 to avoid log(0)

# Replace column names in the OTU table
colnames(log_otu_table) <- column_names

colnames(log_otu_table) <- make.unique(colnames(log_otu_table))

#p2
```
